\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Structure-Preserving Mutations

### 2022-01-13 papers/ast22/main.tex
Rearranged figures.

\begin{algorithm}[b]
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, R, gen}}{
    i $\gets$ 0\;
    TLog, QValid, QDiscarded $\gets$ $\varnothing$\;
    \While{i $<$ N}{
      x $\gets$ Pick(QValid, QDiscarded, gen)\;
      (result, trace) $\gets$ WithTrace(P(x))\;
      \lIf{\KwNot result}{\KwRet Bug(x)}
      \If{Interesting(TLog, trace)}{
        \If{\KwNot Discarded(result)}{
          batch $\gets$ CreateMutationBatch(x, R)\;
          Enqueue(QValid, batch)\;
        }
        \uElseIf{\KwNot Discarded(Parent(x))}{
          batch $\gets$ CreateMutationBatch(x, R)\;
          Enqueue(QDiscarded, batch)\;
        }
      }
      i $\gets$ i+1\;
    }
    \KwRet Ok\;
  }
\caption{\label{algo:mutagen:loop}\mutagen Testing Loop}
\end{algorithm}

@@

\begin{algorithm}[b]
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{Pick}{Pick}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Pick{QValid, QDiscarded, gen}}{
    \If{\KwNot Empty(QValid)}{
      batch $\gets$ Deque(QValid)\;
      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
      \uElse{
        PushFront(QValid, Rest(batch))\;
        \KwRet First(batch)\;
      }
    }
    \If{\KwNot Empty(QDiscarded)}{
      batch $\gets$ Deque(QDiscarded)\;
      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
      \uElse{
        PushFront(QDiscarded, Rest(batch))\;
        \KwRet First(batch)\;
      }
    }
    \lElse{\KwRet Sample(gen)}
  }
\caption{\label{algo:mutagen:pick}\mutagen Seed Selection}
\end{algorithm}

@@

@@

\begin{figure}[b]
%% \vspace{-15pt}
\scriptsize
\newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
\begin{align*}
  \nonumber
  %% \qquad\qquad\qquad\qquad
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
        }
      }
  }
  \right]
  \;
  +
  \;
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{5}}}
        }
      }
  }
  \;
  =
  \;
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\color{red} \small \texttt{1}}
      child {node {\color{red} \small \texttt{2}}
        child {node {\color{red} \small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\color{blue} \small \texttt{5}}}
        }
      }
  }
  \right]
  ,
  \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
  %% \end{equation}
  %% \begin{equation}
  %% \nonumber
  %% \\
  %% \qquad
  \quad
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\small \texttt{5}}}
        }
      }
  }
  \right]
  \;
  +
  \;
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{6}}
          child {node {\small \texttt{7}}
          }
        }
      }
  }
  \;
  =
  \;
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.75cm}
  ]{
    \node {\color{red} \small \texttt{1}}
      child {node {\color{red} \small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\small \texttt{5}}}
        }
        child {node {\color{blue}\small \texttt{6}}
          child {node {\color{blue}\small \texttt{7}}}
        }
      }
  }
  \right]
  ,
  \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
\end{align*}
\vspace{-10pt}
\caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
  internal trace log (represented using brackets).}
\end{figure}
\begin{algorithm}[t]
  \SetKwComment{Comment}{/* }{}
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, R, gen}}{
    $\cdots$\;
    x $\gets$ Pick(QValid, QDiscarded, gen)\;
    (result, trace) $\gets$ WithTrace(P(x))\;
    $\cdots$\;
    \If{Interesting(TLog, trace)}{
      \If{\KwNot Discarded(result)}{
        batch $\gets$ CreateMutationBatch(x, R)\;
        {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
        {\color{red} PushFront(QValid, prio, batch)}\;          %% NEW
      }
      $\cdots$\;
    }
  }
  %% \vspace{5pt}
  \SetKwFunction{Pick}{Pick}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Pick{QValid, QDiscarded, gen}}{
    \If{\KwNot Empty(QValid)}{
      {\color{red} (batch, prio) $\gets$ DequeMin(QValid)}\; %%NEW
      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
      \uElse{
        {\color{red} PushFront(QValid, prio, Rest(batch))}\; %% NEW
        \KwRet First(batch)\;
      }
    }
    \If{\KwNot Empty(QDiscarded)}{\Comment {Analogous to the case above */} }
    $\cdots$\;
  }

>>>>

\begin{algorithm}[t]
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, R, gen}}{
    i $\gets$ 0\;
    TLog, QValid, QDiscarded $\gets$ $\varnothing$\;
    \While{i $<$ N}{
      x $\gets$ Pick(QValid, QDiscarded, gen)\;
      (result, trace) $\gets$ WithTrace(P(x))\;
      \lIf{\KwNot result}{\KwRet Bug(x)}
      \If{Interesting(TLog, trace)}{
        \If{\KwNot Discarded(result)}{
          batch $\gets$ CreateMutationBatch(x, R)\;
          Enqueue(QValid, batch)\;
        }
        \uElseIf{\KwNot Discarded(Parent(x))}{
          batch $\gets$ CreateMutationBatch(x, R)\;
          Enqueue(QDiscarded, batch)\;
        }
      }
      i $\gets$ i+1\;
    }
    \KwRet Ok\;
  }
\caption{\label{algo:mutagen:loop}\mutagen Testing Loop}
\end{algorithm}

@@

@@

\begin{algorithm}[t]
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{Pick}{Pick}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Pick{QValid, QDiscarded, gen}}{
    \If{\KwNot Empty(QValid)}{
      batch $\gets$ Deque(QValid)\;
      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
      \uElse{
        PushFront(QValid, Rest(batch))\;
        \KwRet First(batch)\;
      }
    }
    \If{\KwNot Empty(QDiscarded)}{
      batch $\gets$ Deque(QDiscarded)\;
      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
      \uElse{
        PushFront(QDiscarded, Rest(batch))\;
        \KwRet First(batch)\;
      }
    }
    \lElse{\KwRet Sample(gen)}
  }
  \caption{\label{algo:mutagen:pick}\mutagen Seed Selection}
\end{algorithm}

@@

\begin{figure}[b]
%% \vspace{-15pt}
\scriptsize
\newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
\begin{align*}
  \nonumber
  %% \qquad\qquad\qquad\qquad
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
        }
      }
  }
  \right]
  \;
  +
  \;
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{5}}}
        }
      }
  }
  \;
  =
  \;
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\color{red} \small \texttt{1}}
      child {node {\color{red} \small \texttt{2}}
        child {node {\color{red} \small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\color{blue} \small \texttt{5}}}
        }
      }
  }
  \right]
  %
  \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
  %% \end{equation}
  %% \begin{equation}
  %% \nonumber
  %% \\
  %% \qquad
  \quad
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\small \texttt{5}}}
        }
      }
  }
  \right]
  \;
  +
  \;
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\small \texttt{1}}
      child {node {\small \texttt{2}}
        child {node {\small \texttt{6}}
          child {node {\small \texttt{7}}
          }
        }
      }
  }
  \;
  =
  \;
  \left[
  \tikz[
    baseline=-8mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.75cm}
  ]{
    \node {\color{red} \small \texttt{1}}
      child {node {\color{red} \small \texttt{2}}
        child {node {\small \texttt{3}}
          child {node {\small \texttt{4}}}
          child {node {\small \texttt{5}}}
        }
        child {node {\color{blue}\small \texttt{6}}
          child {node {\color{blue}\small \texttt{7}}}
        }
      }
  }
  \right]
  %
  \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
\end{align*}
\vspace{-10pt}
\caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
  internal execution trace log (represented using brackets).}
\end{figure}

\begin{algorithm}[t]
  \SetKwComment{Comment}{/* }{}
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, R, gen}}{
    $\cdots$\;
    x $\gets$ Pick(QValid, QDiscarded, gen)\;
    (result, trace) $\gets$ WithTrace(P(x))\;
    $\cdots$\;
    \If{Interesting(TLog, trace)}{
      \If{\KwNot Discarded(result)}{
        batch $\gets$ CreateMutationBatch(x, R)\;
        {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
        {\color{red} PushFront(QValid, prio, batch)}\;          %% NEW
      }
      $\cdots$\;
    }
  }
  %% \vspace{5pt}
  \SetKwFunction{Pick}{Pick}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Pick{QValid, QDiscarded, gen}}{
    \If{\KwNot Empty(QValid)}{
      {\color{red} (batch, prio) $\gets$ DequeMin(QValid)}\; %%NEW
      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
      \uElse{
        {\color{red} PushFront(QValid, prio, Rest(batch))}\; %% NEW
        \KwRet First(batch)\;
      }
    }
    \If{\KwNot Empty(QDiscarded)}{\Comment {Analogous to the case above */} }
    $\cdots$\;
  }
